Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-11-06 mprop/mprop.tex
Literature survey on IRL

% TODO: where does the lower bound come from?
Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the reward function $r$. While
the previous paper that considered this IRL model computed maximum likelihood
estimates for $\bm\lambda$ and $\mathbf{u}$, and made an assumption that
$\mathbf{r}$ in \eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11},
we aim to avoid this assumption and use variational inference to approximate the
full posterior distribution $\posterior$. \emph{Variational inference} (VI) is
an approximation technique for probability densities \cite{blei2017variational}.
Let $\approximation$ be our approximating family of probability distributions
for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up to
VI algorithms to optimise $\bm\nu$ in order to minimise the
\emph{Kullback-Leibler} (KL) divergence between the original probability
distribution and our approximation.  KL divergence (asymmetrically) measures how
different the two distributions are and in this case can be defined as follows
\cite{blei2017variational}:
\[ \DKL(\approximation || \posterior) = \Eq \left[ \log
  \frac{\approximation}{\posterior} \right]. \]
Since KL divergence is typically hard to compute, instead of minimising it, VI
typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

By considering full probability distributions instead of point estimates,---as
long as the approximations are able to capture important features of the
posterior---our predictions are likely to be more accurate and rely on fewer
assumptions. Moreover, we hope to make use of various recent advancements in VI
for both time complexity and approximation distribution fit (see Section
\ref{literature}), making the resulting algorithm competitive both in terms of
speed and model fit.

@@

% TODO: review IRL papers, their weaknesses

@@

\subsection{Derivatives}
% TODO: integrals always mean "on all range"
% TODO: some kind of intro

@@

\section{Work Plan}

show how you plan to organize your work, identifying intermediate deliverables
and dates.

1. prove all applications of the dominated convergence theorem

2. develop the algorithm

3. test in on some example problems, drawing some pretty graphs

4. integrate a few extra improvements from various papers

>>>>

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
variational inference to approximate the full posterior distribution
$\posterior$. \emph{Variational inference} (VI) is an approximation technique
for probability densities \cite{blei2017variational}. Let $\approximation$ be
our approximating family of probability distributions for $\posterior$ with its
own hyperparameter vector $\bm\nu$. Then the job of VI algorithms is to optimise
$\bm\nu$ in order to minimise the \emph{Kullback-Leibler} (KL) divergence
between the original probability distribution and our approximation.  KL
divergence (asymmetrically) measures how different the two distributions are,
and in this case can be defined as \cite{blei2017variational}
\[ \begin{split}
    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
    \log\posterior ] \\
    &= \Eq[\log\approximation - \log\pfull] \\
    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
  \end{split}
\]
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as Lebesgue integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

By considering full probability distributions instead of point estimates---as
long as the approximations are able to capture important features of the
posterior---our predictions are likely to be more accurate and rely on fewer
assumptions. Moreover, we hope to make use of various recent advancements in VI
for both time complexity and approximation distribution fit (see Section
\ref{literature}), making the resulting algorithm competitive both in terms of
running time and model fit.

@@

As mentioned in the introduction, most IRL algorithms assume that the reward
function can be represented as a linear combination of features. This assumption
originated in one of the earliest papers on the topic by Ng and Russell
\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
approaches to the problem. The authors also noticed that often
multiple reward functions can explain the same behaviour, and suggested
heuristics for reward functions that are ``far away'' from reward functions that
do not fit the data.

A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
algorithm for the same formulation of the problem with a guarantee to converge
quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
in Abbeel and Ng's approach: the algorithm requires features to be
``appropriately'' scaled, and optimal scaling may not be known. Instead, they
suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
on recovering the reward function).

Ramachandran and Amir \cite{DBLP:conf/ijcai/RamachandranA07} were the first to
formulate IRL in terms of Bayesian learning. While the model is easily
interpretable and able to handle experts that make mistakes, the algorithm can
only handle small state spaces and requires Monte Carlo Markov Chain (MCMC)
sampling for inference.

Ziebart et al. \cite{ziebart2008maximum} keep the linearity assumption, but
introduce an influential idea: resolving the ambiguity when multiple reward
functions explain the data by appealing to the maximum entropy principle.

Choi and Kim \cite{DBLP:conf/ijcai/ChoiK13} extend the Bayesian model to learn
good features as well as the reward function, trying to overcome the limitation
of linearity. However, the approach is quite limiting: all features are assumed
to have Boolean values, and the algorithm simply learns their conjunctions.

Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
learn nonlinear reward functions without harsh restrictions on the problem
domain by using GPs. We base our work primarily on their paper, and the
weaknesses we hope to address have already been covered in the previous section.
A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
aims to harness the power of deep learning by using several layers of GPs,
making the model less dependent on being provided good features. They also use
VI, but with a few simplifying assumptions: deterministic training conditional
for the reward vector, and fully independent training conditional for the latent
state (see Section \ref{sec:vi_literature} for more details).

%A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}
%IRL for large state spaces (haven't read) \cite{DBLP:journals/jmlr/BoulariasKP11}
%Multiple reward functions \cite{DBLP:conf/nips/ChoiK12}
%A recent survey \cite{DBLP:journals/corr/abs-1806-06877}
%EM clustering \cite{DBLP:conf/icml/BabesMLS11}
%Structured priors \cite{DBLP:conf/ewrl/DimitrakakisR11}

\subsection{Variational Inference} \label{sec:vi_literature}

@@

\subsection{Derivatives}
% TODO: some kind of intro

@@

\section{Work Plan}
%show how you plan to organize your work, identifying intermediate deliverables
%and dates.

\begin{enumerate}
\item Prove all applications of the dominated convergence theorem.
\item Implement the algorithm as a more specialised version of Algorithm 2 in
  the black box VI paper.
\item <Evaluation>
\item Experiment with various improvements mentioned in the black box VI paper,
  other VI papers, especially the ones about GPs.
\item Write/finalise the paper.
\end{enumerate}

\subsection{Evaluation}

In order to evaluate our approach, we compare it to two related alternatives:
the original GP IRL paper \cite{DBLP:conf/nips/LevinePK11} and the deep GP
extension \cite{DBLP:conf/uai/JinDAS17}. To the best of our knowledge,... TODO: continue
Against what? GPIRL and DGP-IRL.

For what problems? GPIRL paper has a big toolkit, but with MATLAB.
\begin{itemize}
\item Object world. Fix one, vary the other: number of colours, number of
  examples.
\item Highway driving behaviour: both synthetic and human performance.
\end{itemize}

Using what metrics?
\begin{itemize}
\item Visualise the (true and learned) rewards as shades of grey.
\item Also measure expected value difference in four situations:
  \begin{itemize}
  \item discrete features
  \item discrete feature transfer
  \item continuous features
  \item continuous feature transfer
  \end{itemize}
\item Expected value difference over time rather than #examples (since VI tends
  to be faster than MCMC).
\end{itemize}
